Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2019-03-28 paper.tex
fixes to content, bugs in figure display

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, it remains that current neuroimaging data processing engines do 
    not employ strategies adopted by Big Data processing engines to improve 
    processing time. Here, we 
    evaluate three Big Data processing stategies (in-memory computing, 
    data-locality and lazy-evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast the various 
    processing strategies by using Apache Spark and Nipype as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data/compute ratio of the 
    application to the filesystem/concurrent processes ratio. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Lazy evaluation strategies
were found to increase the likelihood of cache hits, further improving processing time. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone will
not speed-up current 
functional MRI analyses unless coupled with data 
locality and processing around 280 subjects concurrently. 
    In addition, we observe that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of 
an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

Big Data processing engines reduce data-related overheads through 
use of concepts such as data locality, in-memory computing, and lazy evaluation.
Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
framework, execute tasks within the nodes which store the data rather than 
performing costly data transfers across the network. In-memory 
computing, adopted by MapReduce's successor Apache 
Spark~\cite{zaharia2016apache}, stores data in memory whenever 
possible, reducing costly disk I/Os between pipeline steps. 
Lazy evaluation, also used in Spark, enables further Big Data performance 
optimizations, by evaluating results only when necessary. Frameworks 
such as
MapReduce and Spark have become mainstream tools for data analytics, 
although many others, such as 
Dask~\cite{rocklin2015dask}, are emerging. 
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
Nevertheless, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

Scientific applications differ from typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (e.g. Web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}) implemented in consistent software 
libraries, scientific applications often involve 
binary data such as images and signals, processed by a sequence of 
command-line/containerized tools
using a mix of programming languages (C, Fortran, Python, shell 
scripts), referred to as workflows or pipelines. Infrastructure-wise, 
Big Data applications commonly run on 
clouds or dedicated commodity clusters with locality-aware file systems 
such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
usually deployed on large, shared clusters where data is moved between
data nodes and compute nodes using shared file systems such 
as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools.

@@

Our primary field of interest, Neuroimaging, is no exception to the 
generalized rise of data volumes in science due to the joint increase 
of image resolution and subject cohort sizes~\cite{van2014human}. 
Processing engines have been developed with neuroinformatics 
applications in mind, for instance Nipype~\cite{gorgolewski2011nipype} 
or the Pipeline System for Octave and Matlab 
(PSOM~\cite{bellec2012pipeline}). Big Data engines have been used for 
neuroimaging applications too, including in the Thunder 
project~\cite{freeman2014mapping} and in more specific works such 
as~\cite{makkie2019fast}. However, no quantitative performance 
evaluation has been conducted on neuroimaging applications to assess the 
added-value of Big Data engines compared to traditional processing engines.

@@

%\todo{Perhaps talk about modeling and simulation here.}

@@

Apache Spark is a well-established Scala-based processing framework for Big Data, with
APIs in Java, Python and R. Its 
generalized nature allows Spark to not only be applied to batch workflows,
but also SQL queries, iterative machine learning applications, 
data streaming and graph processing. Spark's
main features include data-locality, in-memory processing and lazy evaluation,
which it achieves through its principle abstraction, the Resilient Distributed 
Dataset (RDD).

An RDD is an immutable parallel data structure that achieves fault-tolerance 
through the concept of lineage~\cite{zaharia2010spark}. Rather than permitting
fine-grained transformations, only coarse-grained transformations (applying to
many elements in the RDD), can be applied, thereby making it simple to maintain a 
log of data modifications. This log, known as the lineage, is used
to reproduce any lost data.

Two types of operations can be performed on RDDs:
transformations and actions. Applying a transformation to an RDD produces a new
child RDD through a narrow or wide dependency. A narrow dependency signifies 
that the child is only dependent on a single parent partition, whereas a child 
RDD is dependent on all parent partitions in a wide dependency. Examples of 
Spark transformations include map, filter and join. To materialize an RDD, an
action must be performed, such as a reduce or a collect. Lazy evaluation is 
represented in Spark through the use of transformations and actions. A series of
transformations may be defined without the data ever being materialized. Using 
this strategy, Spark can optimize data processing
throughout the application.

@@

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur within a single application, Spark implements a
Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.

@@

Three different types of schedulers are compatible with Spark: 1) Spark 
Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
scheduler is the default scheduler, whereas YARN is scheduler designed for Hadoop~\cite{white2012hadoop}
clusters and is prepackaged with Hadoop installations. In contrast to Standalone and YARN, 
Mesos was designed to be used in multi-tenant cluster environments. In our experiments,
we focus on the Spark Standalone cluster.

@@

% i use 'popular' for both... widely-used (for Spark)?
Nipype is a popular Python neuroimaging processing engine. It aims at being a 
solution for easily creating reproducible neuroimaging workflows. Although Nipype
does not employ any Big Data processing strategies, it provides access to numerous
plugins to schedulers found in most clusters readily available to researchers, 
such as the Sun/Oracle Grid Engine (SGE/OGE), \href{http://www.adaptivecomputing.com/products/torque/}{TORQUE},
\href{https://slurm.schedmd.com/}{Slurm} and \href{https://research.cs.wisc.edu/htcondor/}{HTCondor}. It also includes its own scheduler, MultiProc,
for parallel processing on single nodes. Nipype additionally provides many built-in 
interfaces to commonly used neuroimaging tools that can be incorporated within the 
workflows. Nipype's ability to easily parallelize workflows in researcher-available 
cluster setups, capture detailed provenance information necessary for reproducibility,
and allow users to easily integrate existing neuroimaging tools, make it preferable
to existing Big Data solutions, which would necessitate modifications to achieve this.

Jobs, or Interfaces, in Nipype, are encapsulated by a Node. A Node dictates that
the job will execute on a single input. However, the MapNode, 
a child variant of the Node, can execute on multiple inputs. All tasks in
Nipype execute in their own uniquely named subdirectory which facilitates provenance 
tracking of inputs and outputs, and also enables checkpointing of the workflow.
In the case of Node failure or application modification, only the nodes which have
been modified (verified by hash) or have no successfully completed, are re-executed.

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 what happens when the processed data fills up the available 
 memory. When engine memory is used, the engine must cleanup 
 unused data with some strategy to avoid crashes. When an in-memory 
 file system is used, the kernel starts swapping memory to disk, and 
 the performance becomes that of local disk writes. In our experiments, 
 we will explore the region  when the data consumed by the 
 application approaches the threshold of available memory.

@@

Storing data on local disks obviously enables data locality, as tasks 
executed on the node where their input files reside do not require 
transfer of input or output data. However, in absence of a more specific 
file system such as HDFS to handle 
file replication across computing nodes, data locality comes at the price
of stringent scheduling restrictions, as tasks can only be scheduled to the
single node that contains their input data.

The performance of local disk accesses is strongly dependent on the 
page caching mechanism provided by the Linux kernel, described in    
details in~\cite{love2010linux}. To summarize, data read from disk 
remains cached in memory until evicted by an LRU (Least Recently Used) 
strategy. When a process invokes the \texttt{read()} system call, the 
kernel will return the data directly from memory if the requested data 
lies in the page cache, realizing a cache \emph{hit}. Cache hits drastically speed-up data 
reads, by masking the disk latency and bandwidth behind a 
memory buffer. In effect, page caching provides in-memory computing 
transparently to the processing engine. However, page cache eviction 
strategies currently cannot be controlled by the application, which 
prevents processing engines to anticipate reads by preloading the 
cache. Scheduling strategies might be designed 
to maximize cache hits though. For instance, lazy 
evaluation could result in more cache hits by scheduling data-dependent 
tasks on the same node.

@@

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a data-compute ratio beyond which the page cache becomes 
asymptotically useless. It should be noted that the transient phase 
during which the page cache fills up might last a significant amount of time, 
in particular when $\dot d$ is positive and small. We intentionally do not model the 
transient phase as it requires detailed knowledge of difficult to estimate parameters  
such as the page cache and the initial amount 
of data in it ($d_0$).

@@

We model a shared file system using its global apparent bandwidth 
$\Delta$, shared by all concurrent processes in the cluster. We are 
aware that such a simplistic model does not describe at all the 
intricacies of systems such as Lustre. In particular, metadata 
management, RPC protocol optimizations and storage optimizations are 
all covered under the apparent bandwidth. We do, however, consider the 
effect of page caching in shared file systems too, since in Linux 
writes to network-mounted volumes benefit from this feature too.

@@

 All experiments were executed on 
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell 
 EMC's Zenith cluster}, a Top-500 machine in the 
 \href{https://www.dellemc.com/en-us/solutions/high-performance-computing/HPC-AI-Innovation-Lab.htm}{Dell 
 EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
 experiments, a Spark cluster was started on our 
 Slurm allocation through access to 16 dedicated nodes.
 Each Compute node has Red Hat Enterprise Linux Server release 7.4 (Maipo) 
 as the base operating system with kernel version 3.10.0-693.17.1.el7.x86\_64
 (patched for Spectre/Meltdown). Dell EMC PowerEdge C6420 with dual Intel Xeon
 Gold 6148/F processors (40 cores per node) and 192GB ($12\times16$~GB), 2666 MHz
 memory, serve as the compute nodes. Each compute has a 120GB M.2 SATA SSD as 
 local disk. A Dell HPC Lustre Solution with a raw storage of 960TB is 
 accesible on each compute node through a 100~GB/s Intel OmniPath network. All
 the nodes connect to a director switch in a 1:1 non-blocking topology.
 The apparent write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script in folder 
 benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.

@@

We used BigBrain~\cite{amunts2013bigbrain}, a 75GB 40$\mu$m isotropic 
histological image of a 65-year-old male's brain. The BigBrain was selected due 
to being the only one of its kind, as there does not yet exist any other
higher-resolution image of a human brain. Moreover, there currently exists a 
lack of standardized tools for processing the BigBrain as a consequence 
of its size. To examine the effects processing the BigBrain has on Big Data 
strategies, we partitioned the full $3845\times3015\times3470$ voxel image into 30 ($5\times3\times2$)
chunks, 125 ($5\times5\times5$) chunks and 750 ($5\times15\times10$) chunks. Additionally, the full 
image was also split in half ($769\times603\times347$ voxels) and the half image was 
partitioned into 125 chunks.

Processing large images is only considered to be part of the Big Data problem 
in neuroscience. The other problem is processing large MRI datasets, 
that is, datasets consisting of many small brain images belonging to 
various different subjects. This situation is commonly observed in 
functional MRI (fMRI), where it is becoming increasingly common to 
process data from hundreds of subjects. Although we have not explored 
explicitly the processing of large MRI datasets, the 75GB BigBrain is 
within the size ballpark~\cite{van2014human} of 
MRI datasets commonly processed in today's studies.

Since both small and large datasets may need to be processed using
the same analysis pipeline, we 
examined the effects of the data management strategies on small 
data as well. For this, we selected a 12MB T1W image belonging to 
subject 1 of OpenNeuro's 
\href{https://openneuro.org/datasets/ds000001/versions/00006/file-display/sub-01:anat:sub-01_T1w.nii.gz}{ds000001 
dataset version 6}. In order to be able to split it into 125 chunks, it 
was necessary to zero-pad the image to the dimensions
$165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.

@@

In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the applications. As incrementing the images is not a time consuming process,
we added a sleep delay to the incrementation task to study the effects 
of tasks duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same file system. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would reside on Lustre. 
We chose to perform our initial input/final output on Lustre as local storage is
typically only accessible to a user for the duration of the execution in HPC
environments.

@@

We conducted four experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size for a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data-locality is not normally preserved in 
Nipype (a new Slurm allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality (see: \href{https://github.com/big-data-lab-team/paper-in-mem-locality/blob/master/benchmark\_scripts/run\_benchmarks.py}{run\_benchmarks.py}).
That is, the chunk were split into partitions, and for each partition,
we requested a Slurm allocation to process the entire 
pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
node. This was possible as no communication was required between the 
processed chunks.

@@

For our first incrementation experiment, we investigated the effects of 
Big Data strategies on varying total data size. To achieve this, we 
increased the number of incrementation iterations from 1, 10 and 100 
times. The total data size would then increase from 75GB, at 1 
iteration, to 7,500GB, at 100 iterations. The total number of chunks 
was 125. Chunks all ran concurrently ($\Gamma$=125) and they were 
equally balanced among 15 nodes, leading to 8 or 9 concurrent jobs per 
node ($\gamma$=9). Task duration was fixed at 3.44 seconds.

@@

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, a greater chunk size signifies 
a decrease in parallelization. However, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelization. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

>>>>

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, current neuroimaging data engines have not adopted any strategies
    to address the rise
    in neuroimaging data. Here, we 
    evaluate three Big Data processing strategies (in-memory computing, 
    data locality and lazy evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast these various 
    strategies using Apache Spark and Nipype as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data-write rate of the 
    application to the filesystem-application concurrent write rate. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Lazy evaluation
strategies
were found to increase the likelihood of cache hits, further improving
    processing time. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. 
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

In Big Data applications, there is a significant cost associated with data movement.
Big Data processing engines have addressed these issues by diminishing the amount 
of data movement that occurs during the execution of an application. The 
MapReduce~\cite{dean2008mapreduce} framework popularized the concept of locality-aware
scheduling, which ensured tasks were scheduled nearest to where the data the data was
located (data locality), thereby reducing the number of network transfers in an application.
However, it was found that performance in MapReduce could be improved upon by also
reducing the disk I/O between tasks by keeping data in-memory whenever possible,
and thus, in-memory computing was added to the Apache Spark~\cite{zaharia2016apache} framework.
Frameworks like Apache Spark also included the lazy evaluation strategy, which performs computations
only when invoked. This allows for additional optimizations such as the regrouping of tasks
and only invoking what is absolutely necessary to obtain the desired result.

Frameworks such as
MapReduce and Spark have become mainstream tools for data analytics, however, many
others, such as Dask~\cite{rocklin2015dask}, are emerging.
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
Nevertheless, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

Scientific applications differ from typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (e.g. Web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}) implemented in consistent software 
libraries, scientific applications often involve 
binary data such as images and signals, processed by a sequence of 
command-line/containerized tools
using a mix of programming languages (C, Fortran, Python, shell 
scripts), referred to as workflows or pipelines. With respect to infrastructure, 
Big Data applications commonly run on 
clouds or dedicated commodity clusters with locality-aware file systems 
such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
usually deployed on large, shared clusters where data is transferred between
data and compute nodes through shared file systems such 
as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
infrastructure have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools.

@@

Our primary field of interest, Neuroimaging, is no exception to the 
generalized rise of data volumes in science due to the joint increase 
of image resolution and subject cohort sizes~\cite{van2014human}. 
Processing engines have been developed with neuroinformatics 
applications in mind, for instance Nipype~\cite{gorgolewski2011nipype} 
or the Pipeline System for Octave and Matlab 
(PSOM~\cite{bellec2012pipeline}). Big Data engines have also been used for 
neuroimaging applications, including the Thunder 
project~\cite{freeman2014mapping} and in more specific works such 
as~\cite{makkie2019fast}. However, no quantitative performance 
evaluation has been conducted on neuroimaging applications to assess the 
added-value of Big Data engines compared to traditional processing engines.

@@

@@

Apache Spark is a well-established Scala-based processing framework for Big Data, with
APIs in Java, Python and R. Its 
generalized nature allows Spark to not only be applied to batch workflows,
but also SQL queries, iterative machine learning applications, 
data streaming and graph processing. Spark's
main features include data locality, in-memory processing and lazy evaluation,
which it achieves through its principle abstraction, the Resilient Distributed 
Dataset (RDD).

An RDD is an immutable parallel data structure that achieves fault-tolerance 
through the concept of lineage~\cite{zaharia2010spark}. Rather than permitting
fine-grained transformations, only coarse-grained transformations (applying to
many elements in the RDD), can be applied, thereby making it simple to maintain a 
log of data modifications. This log, known as the lineage, is used
to reproduce any lost data modifications.

Two types of operations can be performed on RDDs:
transformations and actions. Applying a transformation to an RDD produces a new
child RDD through a narrow or wide dependency. A narrow dependency signifies 
that the child is only dependent on a single parent partition, whereas a child 
RDD is dependent on all parent partitions in a wide dependency. Examples of 
transformations include map, filter and join. To materialize an RDD, an
action must be performed, such as a reduce or a collect. Lazy evaluation is 
represented in Spark through the use of transformations and actions. A series of
transformations may be defined without the data ever being materialized. Using 
this strategy, Spark can optimize data processing
throughout the application.

@@

Although Spark uses in-memory computing, it is not necessary for all the data to
fit in memory. Spark will spill any RDD elements that do not fit in memory to disk.
Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur within a single application, Spark implements a
Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching
significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.

@@

Three different types of schedulers are compatible with Spark: 1) Spark 
Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
scheduler is the default scheduler. The YARN scheduler designed for Hadoop~\cite{white2012hadoop}
clusters and is prepackaged with Hadoop installations, whereas in contrast, 
Mesos was designed to be used in multi-tenant cluster environments. In our experiments,
we focus on the Spark Standalone cluster.

@@

Nipype is a popular Python neuroimaging processing engine. It aims at being a 
solution for easily creating reproducible neuroimaging workflows. Although Nipype
does not employ any Big Data processing strategies, it provides plugins to numerous
schedulers found in most clusters readily available to researchers, 
such as the Sun/Oracle Grid Engine (SGE/OGE), \href{http://www.adaptivecomputing.com/products/torque/}{TORQUE},
\href{https://slurm.schedmd.com/}{Slurm} and \href{https://research.cs.wisc.edu/htcondor/}{HTCondor}. It also includes its own scheduler, MultiProc,
for parallel processing on single nodes. Furthermore, Nipype provides many built-in 
interfaces to commonly used neuroimaging tools that can be incorporated within the 
workflows. Nipype's ability to easily parallelize workflows in researcher-available 
cluster setups, capture detailed provenance information necessary for reproducibility,
and allow users to easily integrate existing neuroimaging tools, make it preferable
over existing Big Data solutions, which would necessitate modifications to achieve this.

Jobs, or Interfaces in Nipype, are encapsulated by a Node. A Node dictates that
the job will execute on a single input. However, the MapNode, 
a child variant of the Node, can execute on multiple inputs. All tasks in
Nipype execute in their own uniquely named subdirectory which facilitates provenance 
tracking of inputs and outputs and also enables checkpointing of the workflow.
In the case of Node failure or application modification, only the nodes which have
been modified (verified by hash) or have not successfully completed, are re-executed.

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 what happens when the processed data fills up the available 
 memory. When engine memory is used, the engine must cleanup 
 unused data to avoid crashes. When an in-memory 
 file system is used, the user is responsible ensuring that the filesystem does not
 reach capacity. Should it be necessary to free up additional memory, the kernel will swap filesystem memory to disk,
 and the performance will become that of local disk writes. In our experiments, 
 we will explore the configuration where the data consumed by the 
 application approaches the threshold of available memory.

@@

Storing data on local disks inevitably enables data locality, since data transfers are
not necessary when tasks are executed on the nodes where the data resides.
However, in absence of a more specific 
filesystem such as HDFS to handle 
file replication across computing nodes, data locality comes at the price
of stringent scheduling restrictions, as tasks can only be scheduled to the
single node that contains their input data.

The performance of local disk accesses is strongly dependent on the 
page caching mechanism provided by the Linux kernel, described in    
details in~\cite{love2010linux}. To summarize, data read from disk 
remains cached in memory until evicted by an LRU (Least Recently Used) 
strategy. When a process invokes the \texttt{read()} system call, the 
kernel will return the data directly from memory if the requested data 
lies in the page cache, realizing a cache \emph{hit}. Cache hits drastically speed-up data 
reads, by masking the disk latency and bandwidth behind a 
memory buffer. In effect, page caching provides in-memory computing 
transparently to the processing engine. However, page cache eviction 
strategies currently cannot be controlled by the application, which 
prevents processing engines from anticipating reads by preloading the 
cache. Scheduling strategies might be designed 
to maximize cache hits, however. For instance, lazy 
evaluation could result in more cache hits by scheduling data-dependent 
tasks on the same node.

@@

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a data-write rate beyond which the page cache becomes 
asymptotically useless. It should be noted that the transient phase 
during which the page cache fills up might last a significant amount of time, 
in particular when $\dot d$ is positive and small. We intentionally do not model the 
transient phase as it requires detailed knowledge of difficult to estimate parameters  
such as the page cache size and the initial amount 
of data in it ($d_0$).

@@

We model a shared file system using its global realized bandwidth 
$\Delta$, shared by all concurrent processes in the cluster. We are 
aware that such a simplistic model does not describe at all the 
intricacies of systems such as Lustre. In particular, metadata 
management, RPC protocol optimizations and storage optimizations are 
all covered under the realized bandwidth. We do, however, consider the 
effect of page caching in shared file systems, since in Linux 
writes to network-mounted volumes benefit from this feature too.

@@

 All experiments were executed on 
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell 
 EMC's Zenith cluster}, a Top-500 machine in the 
 \href{https://www.dellemc.com/en-us/solutions/high-performance-computing/HPC-AI-Innovation-Lab.htm}{Dell 
 EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
 experiments, a Spark cluster was started on a
 Slurm allocation comprised of 16 dedicated nodes.
 Each Compute node has Red Hat Enterprise Linux Server release 7.4 (Maipo) 
 as the base operating system with kernel version 3.10.0-693.17.1.el7.x86\_64
 (patched for Spectre/Meltdown). Dell EMC PowerEdge C6420 with dual Intel Xeon
 Gold 6148/F processors (40 cores per node) and 192GB ($12\times16$~GB), 2666 MHz
 memory, serve as the compute nodes. Each compute has a 120GB M.2 SATA SSD as 
 local disk. A Dell HPC Lustre Solution with a raw storage of 960TB is 
 accesible on each compute node through a 100~Gb/s Intel OmniPath network. All
 the nodes connect to a director switch in a 1:1 non-blocking topology.
 The realized write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script).
 They are reported in Table~\ref{table:bdwdths}.

@@

We used BigBrain~\cite{amunts2013bigbrain}, a 75GB 40$\mu$m isotropic 
histological image of a 65-year-old human brain. The BigBrain was selected due 
to its uniqueness, as there does not yet exist a
higher-resolution image of a human brain. Moreover, there currently exists a 
lack of standardized tools for processing the BigBrain as a consequence 
of its size. To examine the effects processing the BigBrain has on Big Data 
strategies, we partitioned the full $3845\times3015\times3470$ voxel image into 30 ($5\times3\times2$)
chunks, 125 ($5\times5\times5$) chunks and 750 ($5\times15\times10$) chunks. Additionally, the full 
image was also split in half ($769\times603\times347$ voxels) and the half image was 
partitioned into 125 chunks.

Processing large images is only considered to be part of the Big Data problem 
in neuroscience. The other being the processing large MRI datasets, 
that is, datasets consisting of many small brain images belonging to 
various different subjects. This situation is commonly observed in 
functional MRI (fMRI), where it is becoming increasingly common to 
process data from hundreds of subjects. Although we have not explored 
explicitly the processing of large MRI datasets, the 75GB BigBrain is 
within the size ballpark~\cite{van2014human} of 
MRI datasets commonly processed in today's studies.

Since both small and large datasets may need to be processed using
the same analysis pipeline, we 
examined the effects of the data management strategies on small 
data as well. For this, we selected a 12MB T1W image belonging to 
subject 1 of OpenNeuro's 
\href{https://openneuro.org/datasets/ds000001/versions/00006/file-display/sub-01:anat:sub-01_T1w.nii.gz}{ds000001 
dataset version 6}. In order to split the image into 125 equal-sized chunks, it 
was necessary to zero-pad the image to the dimensions
$165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.

@@

To effectively investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consists exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
each task. As incrementing images is not a time consuming process,
we added a sleep delay to the tasks to study the effects 
of tasks duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same file system. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would reside on Lustre. 
We chose to perform our initial input/final output on Lustre as local storage is
typically only accessible to a user for the duration of the execution in HPC
environments.

@@

We conducted four experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size given a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data locality is not normally preserved in 
Nipype (a new Slurm allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality (see: \href{https://github.com/big-data-lab-team/paper-in-mem-locality/blob/master/benchmark\_scripts/run\_benchmarks.py}{run\_benchmarks.py}).
That is, the chunks were split into partitions, and for each partition,
we requested a Slurm allocation to process the entire 
pipeline in parallel using Nipype's MultiProc scheduler on a given 
node. This was possible as no communication was required between the 
processed chunks.

@@

For our first incrementation experiment, we investigated the effects of 
Big Data strategies on varying total data size. To achieve this, we 
increased the number of incrementation iterations from 1, 10 and 100 
times. The total data size would then increase from 75GB, at 1 
iteration, to 7,500GB, at 100 iterations. The total number of chunks 
was 125. Chunks all ran concurrently ($\Gamma$=125) and were 
equally balanced among 15 nodes, leading to 8 or 9 concurrent jobs per 
node ($\gamma$=9). Task duration was fixed at 3.44 seconds.

@@

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, a greater chunk size signifies 
a decrease in parallelism. However, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the data-write rate was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelism. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

For our fourth and final incrementation experiment, we investigated the effects 
of the strategies on different image sizes. We selected the 75GB BigBrain, the
38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
chunks was fixed to 125. Similarly to the previous experiment, the 
total sequential compute time was fixed (10 iterations, 1.76 seconds 
per task), however, due to varying size in total data processed (D), 
the data-write rate varied. Once again, we ensured that the data-write rate fell in 
multiple different regions of the inequality. The data-write rate ranged from 
349.1 MB/s for BigBrain and 174.6MB/s for half BigBrain, to 0.06MB/s for the MRI 
image. Only the 0.06MB/s MRI satisfied the inequality for both Lustre and local 
disk.
